---
title: "RWorksheet_Condag#4c"
author: "Angel Blase"
date: "2024-11-02"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Use the dataset mpg

a. Show your solutions on how to import a csv file into the environment.
```{r}
mpg <- read.csv("mpg.csv")
```

b. Among the 11 variables in the mpg dataset from the ggplot2 package, the categorical variables include manufacturer, model, year, cyl, trans, drv, fl, and class.


c. While the continuous variables are displ, hwy, and cty.


2. Which manufacturer has the most models in this data set? Which model has the most variations?
Show your answer.
```{r}
library(ggplot2)
library(dplyr)

data(mpg)

manufacturer_models <- summarise(group_by(mpg, manufacturer), unique_models = n_distinct(model))
most_models <- arrange(manufacturer_models, desc(unique_models))
most_models <- slice(most_models, 1)
print(most_models)

model_variations <- summarise(group_by(mpg, model), variations = n())
most_variations <- arrange(model_variations, desc(variations))
most_variations <- slice(most_variations, 1)
print(most_variations)

```

a. Group the manufacturers and find the unique models. Show your codes and result.
```{r}
manufacturer_models <- summarise(group_by(mpg, manufacturer), unique_models = n_distinct(model))
manufacturer_models
```

b. Graph the result by using plot() and ggplot(). Write the codes and its result.
```{r}
# Using plot()

manufacturer_models$manufacturer <- as.factor(manufacturer_models$manufacturer)

plot(manufacturer_models$manufacturer, manufacturer_models$unique_models, 
     main = "Unique Models per Manufacturer", 
     xlab = "Manufacturer", 
     ylab = "Number of Unique Models", 
     col = "pink", 
     pch = 19)
```
```{r}
# Using ggplot()

ggplot(manufacturer_models, aes(x = manufacturer, y = unique_models)) +
  geom_bar(stat = "identity", fill = "pink") +
  labs(title = "Unique Models per Manufacturer", 
       x = "Manufacturer", 
       y = "Number of Unique Models") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

2. Same dataset will be used. You are going to show the relationship of the modeland the manufacturer.

a. What does ggplot(mpg, aes(model, manufacturer)) + geom_point() show?
```{r}
ggplot(mpg, aes(model, manufacturer)) + geom_point()
```
Using the code ggplot(mpg, aes(model, manufacturer)) + geom_point() creates a scatter plot with model on the x-axis and manufacturer on the y-axis. Each point on the plot represents a specific car model from the mpg dataset, plotted according to its associated manufacturer.


b. The initial scatter plot created with ggplot(mpg, aes(model, manufacturer)) + geom_point() is not very useful because it displays a lot of overlapping points, making it hard to distinguish between different car models and their manufacturers. This clutter can obscure patterns and relationships in the data. To make the plot more informative, it could be enhanced by adding color to represent another variable, like the car class, which would provide more context. Additionally, adjusting the size of the points based on factors like highway miles per gallon would help convey more information.


3. Plot the model and the year using ggplot(). Use only the top 20 observations. Write the codes and
its results.
```{r}
top_20_models <- head(mpg, 20)

ggplot(top_20_models, aes(x = model, y = year)) +
  geom_point(color = "purple", size = 3) +
  labs(title = "Top 20 Car Models by Year",
       x = "Car Model",
       y = "Year") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

