---
title: "RWorksheet_Condag#4b"
author: "Angel Blase"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Using Loop Function
1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
```{r}
vectorA <- c(1, 2, 3, 4, 5)
x <- matrix(vectorA, nrow = 5, ncol = 5)

for (i in 1:5) {
  for(j in 1:5) {
    x[i, j] <- abs(i - j)
  }
}

x
```
2. Print the string "*" using for() function. The output should be the same as shown in Figure.
```{r}
for(i in 1:5) {
  line <- rep('"*"', i)
  cat(line, sep = " ")
  cat("\n")
}
```
3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
#start <- as.integer(readline("Enter the starting term for the Fibonacci Sequence: "))
start <- 3
x <- 0
y <- 1
index <- 1

repeat {
  if(index >= start) {
    if (x > 500) {
      break
    }
    cat(x, " ")
  }
  
  nxt <- x + y
  x <- y
  y <- nxt
  index <- index + 1
}

cat("\n")
```
# Using Basic Graphics (plot(),barplot(),pie(),hist())

4. Import the dataset

a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result.
```{r}
shoeSize <- read.csv("Shoe_sizes.csv")
head(shoeSize)
```

b. Create a subset for gender(female and male). How many observations are there in Male? How about in Female? Write the R scripts and its output.
```{r}
male <- subset(shoeSize, Gender == "M")
female <- subset(shoeSize, Gender == "F")

n_males <- nrow(male)
n_females <- nrow(female)

n_males
n_females
```

c. Create a graph for the number of males and females for Household Data. Use plot(),chart type = barplot. Make sure to place title, legends, and colors. Write the R scripts and its result.
```{r}
genderCounts <- table(shoeSize$Gender)

barplot(genderCounts,
        main = "Number of Males and Females",
        xlab = "Gender",
        ylab = "Counts",
        col = c("purple", "pink"),
        legend = row.names(genderCounts)
        )
```

5. The monthly income of Dela Cruz family was spent on the following:

a. Create a piechart that will include labels in percentage.Add some colors and title of the chart. Write the R scripts and show its output.
```{r}
expenses <- c(Food = 60, Electricity = 10, Savings = 5, Miscellaneous = 25)
percentages <- round(expenses / sum(expenses) * 100)
labels <- paste(names(expenses), percentages, "%")
colors <- c("pink", "purple", "yellow", "skyblue")
pie(expenses,
    labels = labels,
    col = colors,
    main = "Dela Cruz Family Monthly Expenses")
```

6. Use the iris dataset.
data(iris)

a. Check for the structure of the dataset using the str() function. Describe what you have seen in the output.
```{r}
data(iris)
str(iris)
```
Description: The image shows the output of the str(iris) command in R, which provides a summary of the iris dataset. This dataset has 150 rows and 5 columns. The columns include Sepal.Length and Sepal.Width, which represent the length and width of sepals, and Petal.Length and Petal.Width, which represent the length and width of petals. All these columns are numeric. The fifth column, Species, is a categorical (factor) variable with three levels: "setosa," "versicolor," and "virginica," representing different iris species. The summary displays sample values for each column to give an overview of the data structure.


b. Create an R object that will contain the mean of the sepal.length,
sepal.width,petal.length,and petal.width. What is the R script and its result?
```{r}
mean_values <- colMeans(iris[, c("Sepal.Length", "Sepal.Width", "Petal.Length", "Petal.Width")])
mean_values
```

c. Create a pie chart for the Species distribution. Add title, legends, and colors. Write the R script and its result.
```{r}
species_counts <- table(iris$Species)

pie(species_counts, 
    main = "Species Distribution in Iris Dataset",
    col = c("skyblue", "pink", "lightgreen"),
    labels = paste(names(species_counts), "-", species_counts))

legend("topright", 
       legend = names(species_counts), 
       fill = c("skyblue", "pink", "lightgreen"), 
       title = "Species"
)

```

d. Subset the species into setosa, versicolor, and virginica. Write the R scripts and show the last six (6) rows of each species.
```{r}
setosa <- subset(iris, Species == "setosa")
versicolor <- subset(iris, Species == "versicolor")
virginica <- subset(iris, Species == "virginica")

cat("Last 6 rows of Setosa:\n")
print(tail(setosa, 6))

cat("\nLast 6 rows of Versicolor:\n")
print(tail(versicolor, 6))

cat("\nLast 6 rows of Virginica:\n")
print(tail(virginica, 6))
```

e. Create a scatterplot of the sepal.length and sepal.width using the different species(setosa,versicolor,virginica). Add a title = “Iris Dataset”, subtitle = “Sepal width and length, labels for the x and y axis, the pch symbol and colors should be based on the species.
```{r}
species_colors <- c("setosa" = "red", "versicolor" = "green", "virginica" = "blue")
species_pch <- c("setosa" = 16, "versicolor" = 17, "virginica" = 18)

plot(iris$Sepal.Length, iris$Sepal.Width,
     main = "Iris Dataset",
     sub = "Sepal width and length",
     xlab = "Sepal Length",
     ylab = "Sepal Width",
     col = species_colors[iris$Species],
     pch = species_pch[iris$Species])

legend("topright", legend = levels(iris$Species), 
       col = species_colors, pch = species_pch, 
       title = "Species")
```

f. Interpretation: The scatterplot shows the relationship between sepal length and sepal width for the three species in the iris dataset: Setosa, Versicolor, and Virginica. Each species is represented by a different color and symbol, making it easy to see how they differ. Setosa (red points) generally has a wider sepal width compared to its sepal length and is well-separated from the other two species. Versicolor (green points) and Virginica (blue points) have more overlap in their sepal sizes, but Virginica tends to have longer sepals. This pattern suggests that Setosa is more distinct in terms of sepal measurements, while Versicolor and Virginica are more similar.

# Basic Cleaning and Transformation of Objects

7. 